### 2017-02-10 draft-md-ct.tex
steel plate covering shame, and FTRL = MD
For the time being needs to compile using:  LaTeX + dvips + ps2pdf
Until the steel plate (joke) is removed from the road :-)

% ========== Pstricks
%\usepackage{pst-all}

@@

\noindent
Let's put 
$\displaystyle y(t) = \int_{0}^{t} \dot{u}(s) ds = -\eta \int_{0}^{t} \nabla\ell_{s}(x(s)) ds$.
Then for a fixed $x$,
\begin{equation}
\int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), x}
= \frac{\ab{y(t), x}}{\eta}
\leq \frac{R^{*}(y(t)) + R(x)}{\eta}\,.
\label{eq.using.fenchel.ineq}
\end{equation}
But noting that $x(t) = \nabla R^{*}(y(t))$, 
which means that $x(t)$ is the optimizer in the definition of $R^{*}$, then
\begin{align}
\frac{R^{*}(y(t))}{\eta}
&= \frac{\ab{y(t), x(t)} - R(x(t))}{\eta} \notag \\[1mm]
&= \int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), x(t)} ds - \frac{R(x(t))}{\eta}\,.
\label{eq.something}
\end{align}
Consider the function $\varphi(t,x) = \int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), x} ds$.
The time derivative is
$\dot{\varphi}(t,x) = \ab{-\nabla\ell_{t}(x(t)), x}$.
In particular, $\dot{\varphi}(t,x(t)) = \ab{-\nabla\ell_{t}(x(t)), x(t)}$,
and integrating (note that $\varphi(0,x(0)) = 0$) we get
$$
\int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), x(t)} ds 
= \varphi(t,x(t)) = \int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), x(s)} ds\,.
$$
% omar's note: this last thing looked like magic to me!
Therefore continuing from \eqref{eq.using.fenchel.ineq} and \eqref{eq.something} above,
and rearranging,
$$
\int_{0}^{t} \ab{\nabla\ell_{s}(x(s)), x(s)} ds - \int_{0}^{t} \ab{\nabla\ell_{s}(x(s)), x} ds
\leq \frac{R(x) - R(x(t))}{\eta}\,.
$$
So using the linearization trick (see \eqref{eq.linearizing.loss} above) we get
$$
\int_{0}^{T} [\ell_{t}(x(t)) - \ell_{t}(x^{*})] dt
\leq \int_{0}^{T} \ab{\nabla\ell_{t}(x(t)), x(t) - x} dt
\leq \frac{R(x) - R(x(T))}{\eta}\,.
$$

>>>>

% ========== Pstricks
\usepackage{pst-all}

@@

\noindent
Let's put 
$\displaystyle y(t) = \int_{0}^{t} \dot{u}(s) ds = -\eta \int_{0}^{t} \nabla\ell_{s}(x(s)) ds$.
Then for a fixed $x$,
\begin{equation}
\int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), x}
= \frac{\ab{y(t), x}}{\eta}
\leq \frac{R^{*}(y(t)) + R(x)}{\eta}\,.
\label{eq.using.fenchel.ineq}
\end{equation}
But noting that $x(t) = \nabla R^{*}(y(t))$, 
which means that $x(t)$ is the optimizer in the definition of $R^{*}$, then
\begin{align}
\frac{R^{*}(y(t))}{\eta}
&= \frac{\ab{y(t), x(t)} - R(x(t))}{\eta} \notag \\[1mm]
&= \int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), x(t)} ds - \frac{R(x(t))}{\eta}\,.
\label{eq.something}
\end{align}
Consider the function $\varphi(t,x) = \int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), x} ds$.
The time derivative is
$\dot{\varphi}(t,x) = \ab{-\nabla\ell_{t}(x(t)), x}$.
In particular, $\dot{\varphi}(t,x(t)) = \ab{-\nabla\ell_{t}(x(t)), x(t)}$,
and integrating (note that $\varphi(0,x(0)) = 0$) we get
$$
\int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), x(t)} ds 
= \varphi(t,x(t)) = \int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), x(s)} ds\,.
$$
% omar's note: this last thing needs to be fixed!

% Picture 
\psset{unit=1.4pt} \vspace{3mm} \hspace{8cm}
\pspicture(0,0)(0,0)
\pspolygon[fillcolor=gray,fillstyle=crosshatch*](0,0)(70,0) (70,50)(0,50)
\rput(35,25){\psframebox*[framearc=.3]{\parbox[c]{2.5cm}{\footnotesize\centering STEEL PLATE ON ROAD}}}
\endpspicture

% omar's note: the following to be added back when the above is fixed!
%\noindent
%Therefore continuing from \eqref{eq.using.fenchel.ineq} and \eqref{eq.something} above,
%and rearranging,
%$$
%\int_{0}^{t} \ab{\nabla\ell_{s}(x(s)), x(s)} ds - \int_{0}^{t} \ab{\nabla\ell_{s}(x(s)), x} ds
%\leq \frac{R(x) - R(x(t))}{\eta}\,.
%$$
%So using the linearization trick (see \eqref{eq.linearizing.loss} above) we get
%$$
%\int_{0}^{T} [\ell_{t}(x(t)) - \ell_{t}(x^{*})] dt
%\leq \int_{0}^{T} \ab{\nabla\ell_{t}(x(t)), x(t) - x} dt
%\leq \frac{R(x) - R(x(T))}{\eta}\,.
%$$

\section*{Equivalence between FTRL and MD}

Some notes to clarify the equivalence between 
Follow the Regularized Leader (FTRL) and Mirror Descent (MD) algorithms.

The FTRL algorithm (discrete time) uses the initialization and updates
\begin{align}
x_1 &= ??? \label{eq.ftrl1} \tag{1ftrl} \\
x_{t+1} &= \arg\min_{x \in \R^d} \Bigl\{ \eta \Bigab{ \sum_{s=1}^{t} \nabla\ell_s(x_{s}), x} + R(x)\Bigr\}\,.
 \label{eq.ftrl2} 
\tag{2ftrl}
\end{align}
Here $R(x)$ is the regularizer used by FTRL.
To find an explicit form for the optimizer, let
$$
G(x) = \eta \sum_{s=1}^{t} \ab{ \nabla\ell_s(x_{s}), x} + R(x)\,.
$$
Then $\nabla G(x) = \eta \sum_{s=1}^{t} \nabla\ell_s(x_{s}) + \nabla R(x)$
and it is clear that $\nabla G(x) = 0$ 
when $x = \nabla R^{*} \bigl( -\eta \sum_{s=1}^{t} \nabla\ell_s(x_{s}) \bigr)$.
The second derivative $\nabla^2 G(x) = \nabla^2 R(x)$ is positive,
so this point $x$ gives the minimum.
So the update rule \eqref{eq.ftrl2} becomes
$$
x_{t+1} = \nabla R^{*} \bigl( -\eta \sum_{s=1}^{t} \nabla\ell_s(x_{s}) \bigr)\,.
$$
Defining $u_{t+1} = -\eta \sum_{s=1}^{t} \nabla\ell_s(x_{s})$ with $u_{0} = 0$,
we get $u_{t+1} = u_{t} - \eta \nabla\ell_t(x_{t})$,
and the update can be rewritten as the following:
\begin{align*}
u_{t+1} &= u_{t} - \eta \nabla \ell_{t}(x_{t})\,, \\
x_{t+1} &= \nabla R^{*}(u_{t+1})\,. 
\end{align*}
These are clearly the equations \eqref{eq.dtime.1}--\eqref{eq.dtime.2} of MD.

Conversely, starting from the equations \eqref{eq.dtime.1}--\eqref{eq.dtime.2} of MD,
we can revert the process and obtain the equations \eqref{eq.ftrl1}--\eqref{eq.ftrl2} of FTRL.
